---
title: "Enigmatic Compounds (With Political Leaning as Covariate)"
author: "Qi Yu"
date: "2024-03-01"
output: html_document
---

REFERENCE:

[Aggregate function](https://www.guru99.com/r-aggregate-function.html)

[Proportional odds assumption](https://peopleanalytics-regression-book.org/gitbook/ord-reg.html)

[Plotting Likert data](https://jakec007.github.io/2021-06-23-R-likert/)


```{r warning=FALSE,message=FALSE}
library(ordinal)
library(dplyr)
library(tidyr)
library(knitr)
library(Hmisc)
library(likert)
library(janitor)
library(RColorBrewer)
```

# 1. Data preprocessing

```{r}
df <- readRDS("./data/data_with_political_leaning.rds")

df <- mutate(df, 
             prolificID = factor(prolificID),
             item_num_only = factor(item_num_only),
             conditions = factor(conditions),
             response_1 = factor(response_1),
             list = factor(list),
             leftright_mapped = factor(leftright_mapped),
             leftright = factor(leftright)
) 

df$conditions <- relevel(df$conditions, ref="phrasal")

kable(head(df))
```

# 2. Descriptive statistics 

## 2.1 Distribution of participants' ratings

```{r}
df_rate_percent <- df %>%
  tabyl(conditions, response_1) %>%
  adorn_percentages("row") %>% 
  mutate_if(is.numeric, ~ . * 100) %>% # Change to percentage
  mutate_if(is.numeric, round, 0)

df_rate_percent <- as.data.frame(df_rate_percent) 
colnames(df_rate_percent) <- c("Item", "1", "2", "3", "4", "5", "6", "7")
df_rate_percent
```


```{r}
# Get the color order and reverse it: Using blue for positive categories and red for negative categories
colors <- rev(brewer.pal(7, "Spectral"))

p <- plot(likert(summary = df_rate_percent), legend.position="right") +
  scale_fill_manual(values = colors) + 
  theme(text = element_text(size = 21)) +
  guides(fill = guide_legend(title="Responses")) 

p

ggsave(
  "./plots/rating_distribution.jpg",
  plot = p,
  width = 10, height = 3,
  dpi = 300)
```

## 2.2 Distributions of other meta data

```{r results = 'hide'}
df_by_participant <- df %>%
  group_by(prolificID) %>% 
  dplyr::summarize(list = first(list), # Using the "::" syntax: The package Hmisc interferrs with dplyr 
            age = first(age),
            duration = first(time_in_minutes),
            gender = first(gender),
            leftright = first(leftright_mapped)
            )

kable(head(df_by_participant))
```

**Participants per list:**
```{r}
table(df_by_participant$list)
```

**Descriptive statistics of age:**
```{r}
mean(df_by_participant$age)
sd(df_by_participant$age)
```

**Descriptive statistics of gender:**
```{r}
table(df_by_participant$gender)
```

**Descriptive statistics of political leaning:**
```{r}
table(df_by_participant$leftright)
```


# 3. Fit models 

Random effects: ID, item

## 3.1 A null model
* Intercept only 
* Random intercepts only 

```{r}
model0 <- clmm(response_1 ~ 1 + (1|item_num_only) + (1|prolificID), data=df, link="logit", Hess=TRUE)
summary(model0)
```

## 3.2 Model 1: 
* Conditions as fixed effect; 
* Random intercepts only

```{r}
model1 <- clmm(response_1 ~ conditions + (1|item_num_only) + (1|prolificID), data=df, link="logit", Hess=TRUE)
summary(model1)
```

## 3.3 Model 2:
* Conditions as fixed effect; 
* Random intercepts and random intercepts for items; Only random intercepts for participants

```{r}
model2 <- clmm(response_1 ~ conditions + (conditions|item_num_only) + (1|prolificID), data=df, link="logit", Hess=TRUE) # 
summary(model2)
```

## 3.4 Model 3: 
* Conditions as fixed effect; 
* Random intercepts and random intercepts for participants; Only random intercepts for items

```{r}
model3 <- clmm(response_1 ~ conditions + (1|item_num_only) + (conditions|prolificID), data=df, link="logit", Hess=TRUE)
summary(model3)
```

## 3.5 Model 4: 
* Conditions as fixed effect; 
* Random intercepts and random intercepts for both participants and items

```{r}
model4 <- clmm(response_1 ~ conditions + (conditions|item_num_only) + (conditions|prolificID), data=df, link="logit", Hess=TRUE)
summary(model4)
```

## 3.6 Model 5: 
* Conditions and political leanings as fixed effects; 
* Random intercepts and random intercepts for both participants and items

```{r}
model5 <- clmm(response_1 ~ conditions + leftright_mapped + (conditions|item_num_only) + (conditions|prolificID), data=df, link="logit", Hess=TRUE)
summary(model5)
```

## 3.7 Model 6: 
* Conditions and political leanings as fixed effects; 
* An interaction term of conditions and political leanings is included;
* Random intercepts and random intercepts for both participants and items

```{r}
model6 <- clmm(response_1 ~ conditions*leftright_mapped + (conditions|item_num_only) + (conditions|prolificID), data=df, link="logit", Hess=TRUE)
summary(model6)
```


# 4. Compare goodness of fit
```{r}
anova(model0, model1)
```

```{r}
anova(model1, model2)
```

```{r}
anova(model1, model3)
```

```{r}
anova(model2, model4)
```


```{r}
anova(model3, model4)
```

```{r}
anova(model4, model5)
```

```{r}
anova(model5, model6)
```



